---
title: "sbtools -- getting started"
author: "dblodgett@usgs.gov"
date: "2023-11-17"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sbtools -- getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}

local <- (Sys.getenv("BUILD_VIGNETTES") == "TRUE")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6, 
  fig.height=4,
  eval=local
)

oldoption <- options(scipen = 9999)
```

`sbtools` provides access and upload utilities for [ScienceBase](https://sciencebase.gov)

Much of sbtools does not require authentication.

```{r}
## Examples

library(sbtools)

# Query ScienceBase for data about Antarctica
query_sb_text('Antarctica', limit=1)

# Query for a specific DOI
query_sb_doi('10.5066/P92U7ZUT')

# Inspect the contents of the above item
children <- item_list_children('5669a79ee4b08895842a1d47')

sapply(children, function(child) child$title)

item_get(children[[1]])
```
`sbtools` uses an `sbitem` S3 object that is a list containing the `sbjson` associated with the item in question.

```{r}

item <- item_get(children[[1]])

class(item)

class(unclass(item))

```
Much of `sbtools` is intended to be used after authentication. This part of `sbtools` is changing in Winter, 2023-2024 as ScienceBase transitions for password-based authentication to two factor and token based authentication. This next section was rendered with the author's (dblodgett@usgs.gov) private credentials.

The old way used `authenticate_sb()` with a username and password. The password could be cached using the `keyring` package.

```{r}
authenticate_sb(Sys.getenv("sb_user"))

my_home_item <- user_id()

new_item <- item_create(title = 'new test item', parent_id = my_home_item)

test.txt <- file.path(tempdir(), 'test.txt')

writeLines(c('this is','my data file'), test.txt)

item_append_files(new_item, test.txt)

item_list_files(new_item)$fname

item_rm(new_item)

unlink(test.txt)

# restart or clean session to reauthenticate differently
sbtools:::clean_session()
```

The new way uses `initialize_sciencebase_session()` to open a browser window. Once logged in, you can get a token from the user drop down in the upper right. For this example, that token has been saved as an environment variable.

For this to work, a `token_text` environment variable must be available and valid.

```{r}
initialize_sciencebase_session(Sys.getenv("sb_user"), Sys.getenv("token_text"))

my_home_item <- user_id()

new_item <- item_create(title = 'new test item', parent_id = my_home_item)

test.txt <- file.path(tempdir(), 'test.txt')

writeLines(c('this is','my data file'), test.txt)

item_append_files(new_item, test.txt)

item_list_files(new_item)$fname

item_rm(new_item)

unlink(test.txt)
```

```{r teardown, include=FALSE}
options(oldoption)
sbtools:::clean_session()
```